Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest

### 2021-08-26 Manuscript.Rmd
Final version accepted in JVS

---
title: Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/journal-of-vegetation-science.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
library(rstan)
options(mc.cores = 4)
rstan_options(auto_write = T)
library(bayesplot)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data, include = F}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n())
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, trait, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
# ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
#   geom_density(alpha = .5) +
#   facet_wrap(~ strata + trait, scales = "free") +
#   scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
# reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
#   filter(variable != "BA") %>% 
#   ggplot(aes(value, SES, col = strata)) +
#   facet_grid(trait ~ variable, scales = "free") +
#   geom_point() +
#   geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
#   geom_smooth(method = "lm") +
#   theme(legend.position = "bottom", axis.title = element_blank())
```

```{r, eval=F}
filter(trees, community %in% filter(R, Canopy < 20, Slope > 30)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low canopy (<20m) and strong slope (>30)")
filter(trees, community %in% filter(R, Slope < 15)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low slope (<15)")
filter(PFT, strata == "U", trait == "WD") %>% 
  ggplot(aes(traitvalue)) +
  geom_histogram() +
  ggtitle("WD distribution among understory species.")
left_join(trees, PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ strata, scales = "free") +
  ggtitle("Number of trees per community along slope gradient.")
left_join(trees, filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  filter(!is.na(traitvalue)) %>% 
  mutate(WDclass = cut(traitvalue, 
                       breaks = c(0, 0.5, 0.7, 1), 
                       labels = c("low", "mid", "high"))) %>% 
  group_by(WDclass, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ WDclass, scales = "free", nrow = 2) +
  ggtitle("Number of trees per community along slope gradient",
          "depending on WD category (low < 0.5 and high > 0.7).")
```

@@

```{r Uppangala, fig.cap="Digital canopy model built from LiDAR data of the 10-ha forest permanent Uppangala plot , India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
ggplot() +  
geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
    annotation_custom(grob = ggplotGrob(ggplot() +  
                                        geom_polygon(data = fortify(india),
                                                     aes(long, lat, group=group),
                                                     colour = "black", alpha = 0) +
                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
                                                   aes(long, lat), col = "red", size = 3) +
                                        coord_equal() +
                                        cowplot::theme_map() +
                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
                                              plot.background	= element_rect(fill = "white"))), 
                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
```

@@

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See trait abbreviations in Table \\@ref(tab:traits).'}
pca.plot <- autoplot(PFT.pca, 
         data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
         colour = "strata",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  # ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
```

@@

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of leaf area (LA) , wood density (WD), leaf thickness (LT), leaf dry matter content (LDMC) and  specific leaf area (SLA) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** paired Student p-value<0.001).", fig.height=4}
ggplot(GWM, aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
```

>>>>

---
title: Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest
output:
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/journal-of-vegetation-science.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
library(rstan)
options(mc.cores = 4)
rstan_options(auto_write = T)
library(bayesplot)
library(ggmap)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r, eval=FALSE}
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
# Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats(cs = c(30,30))
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- raster::extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/bin/')
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, raster::extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- raster::extract(DEMderiv(Env$DEM, 'wetness', path = '/usr/bin/'),
                                    as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- raster::extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
R <- rownames_to_column(R, "community")
save(R, file = file.path("draft_save", 'R2.Rdata'))
```

```{r data, include = F}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  dplyr::summarise(Abundance = n())
# Matrix R
# load(file.path("draft_save", 'R.Rdata'))
load(file.path("draft_save", 'R2.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
# left_join(GWM, R) %>% 
#   reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
#   filter(variable != "BA") %>% 
#   ggplot(aes(value, GWM, col = strata)) +
#   facet_grid(trait ~ variable, scales = "free") +
#   geom_point() +
#   geom_smooth(method = "lm") +
#   theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, trait, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
# ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
#   geom_density(alpha = .5) +
#   facet_wrap(~ strata + trait, scales = "free") +
#   scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
# reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>%
#   filter(variable != "BA") %>%
#   ggplot(aes(value, SES, col = strata)) +
#   facet_grid(trait ~ variable, scales = "free") +
#   geom_point() +
#   geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
#   geom_smooth(method = "lm") +
#   theme(legend.position = "bottom", axis.title = element_blank())

```

```{r, eval=FALSE}
trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  dplyr::group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
```

```{r, eval=F}
filter(trees, community %in% filter(R, Canopy < 20, Slope > 30)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low canopy (<20m) and strong slope (>30)")
filter(trees, community %in% filter(R, Slope < 15)$community) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  arrange(community, desc(traitvalue)) %>% 
  kable(caption = "Community with low slope (<15)")
filter(PFT, strata == "U", trait == "WD") %>% 
  ggplot(aes(traitvalue)) +
  geom_histogram() +
  ggtitle("WD distribution among understory species.")
left_join(trees, PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ strata, scales = "free") +
  ggtitle("Number of trees per community along slope gradient.")
left_join(trees, filter(PFT, trait == "WD")) %>% 
  filter(strata == "U") %>% 
  filter(!is.na(traitvalue)) %>% 
  mutate(WDclass = cut(traitvalue, 
                       breaks = c(0, 0.5, 0.7, 1), 
                       labels = c("low", "mid", "high"))) %>% 
  group_by(WDclass, community) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(R) %>% 
  ggplot(aes(Slope, Abundance)) +
  geom_point() +
  facet_wrap(~ WDclass, scales = "free", nrow = 2) +
  ggtitle("Number of trees per community along slope gradient",
          "depending on WD category (low < 0.5 and high > 0.7).")

trees %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  left_join(select(PFT,SpCode, strata)) %>% 
  unique() %>% 
  filter(!is.na(strata)) %>% 
  group_by(community, strata) %>% 
  summarise(N = sum(Abundance), S = length(unique(SpCode))) %>% 
  group_by(strata) %>% 
  # summarise(N = sd(N), S = sd(S))
  summarise(N = mean(N), Nsd = sd(N),
            S = mean(S), Ssd = sd(S))
  # group_by(strata, community) %>% summarise(N = sum(Abundance), S = length(unique(SpCode))) %>% group_by(strata) %>% summarise(N = mean(N), S = mean(S))
```

```{r, eval=F}
data <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  mutate(LMA = 1/SLA, LT.LDMC = LT*LDMC) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(mean, na.rm = T)
summary(lm(LMA ~ LT.LDMC, data))
ggplot(data, aes(LT.LDMC, LMA)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("LMA = 0.30 + 0.17 LT x LDMC",
          "r2 = 0.48; n = 267; P < 0.001") +
  xlab("LT x LDMC")

```

@@

```{r Uppangala, fig.cap="Digital canopy model built from LiDAR data of the 10-ha forest permanent Uppangala plot , India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
d <- 12
lng <- 75.662778
lat <- 12.537500
bbox <- c(left = lng-d, bottom = lat-d, right = lng+d, top = lat+d)
map <- get_stamenmap(bbox, zoom = 5)
g <- ggplot() +  
  geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  theme(axis.line = element_blank(), legend.position = c(0.9, 0.5), axis.text = element_text(face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA)) +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
  ggsn::north(symbol = 10, x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 30, y.max = max(DCM$y)-10) +
  xlab("Longitude") + ylab("Latitude") +
  scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0)) +
  annotation_custom(grob = ggplotGrob(
    ggmap(map) +
      geom_point(x = lng, y = lat, colour = "red", size = 4) +
      theme(axis.title = element_blank(), axis.line = element_blank(), 
            axis.ticks = element_blank(), axis.text = element_blank(),
            panel.border = element_rect(fill = NA))
  ),
  xmin = min(DCM$x), xmax = min(DCM$x) + 200,
  ymin = max(DCM$y) - 200, ymax = max(DCM$y))
# ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
ggsave(g, filename = "figures/Figure_1.tiff", device = "tiff", dpi = 600, bg = "white", width = 8, height = 8, compression = "lzw")
```

@@

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See trait abbreviations in Table \\@ref(tab:traits).'}
# pca.plot <- autoplot(PFT.pca, 
#          data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
#          colour = "strata",
#          loadings.label.size = 6,
#          loadings.label.colour = 'black',
#          loadings = T, loadings.label = T, loadings.colour = 'black',
#          loadings.label.vjust = 1.2) +
#   coord_equal() +
#   geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
#   geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
#   xlab('Axis 1') + ylab('Axis 2') +
#   theme(legend.position = c(1, 1), 
#         legend.justification = c(1, 1)) +
#   scale_color_discrete("Guild")
# spectrum.boxplots <- pca.plot$data %>%
  # select(Comp.1, Comp.2, strata) %>%
  # rename(`Axis 1` = Comp.1, `Axis 2` = Comp.2) %>%
  # melt(id.vars = "strata", variable.name = "spectrum") %>%
  # ggplot(aes(strata, value, fill = strata)) +
  # # ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  # geom_boxplot() +
  # facet_wrap(~spectrum, nrow = 2) +
  # scale_fill_discrete("Guild")
g <- autoplot(PFT.pca, 
         data = na.omit(reshape2::dcast(PFT, strata + SpCode ~ trait, value.var = "traitvalue")),
         colour = "strata",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  stat_ellipse(aes(group = strata, col = strata), linetype = "dashed") +
  xlab('Axis 1') + ylab('Axis 2') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1),
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  scale_color_discrete("Guild")
ggsave(g, filename = "figures/Figure_2.tiff", device = "tiff", dpi = 600, bg = "white", width = 8, height = 8, compression = "lzw")
```

@@

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of leaf area (LA) , wood density (WD), leaf thickness (LT), leaf dry matter content (LDMC) and  specific leaf area (SLA) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** paired Student p-value<0.001).", fig.height=4, fig.width=10}
g <- GWM %>% 
  mutate(trait = recode(trait, 
                        "SLA" = "A", 
                        "LDMC" = "B",
                        "LT" = "C", 
                        "LA" = "D",
                        "WD" = "E")) %>% 
  mutate(trait = factor(trait, levels = c("A", "B", "C", "D", "E"))) %>%
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  geom_text(aes(label = label), label.size = 0.2,
            data.frame(
              strata = "C",
              trait = LETTERS[1:5],
              label = c("A SLA", "B LDMC", "C LT", "D LA", "E WD"),
              GWM = c(2.8, 435, 28, 10, 0.7)
            )) +
  facet_wrap(~trait, scales = "free", nrow = 1,
             strip.position = "left", 
             labeller = as_labeller(c(A = "SLA (m²/kg)", B = "LDMC (mg/g)",
                                      C = "LT (µm)", D = "LA (mm²)", E = "WD (g/cm³)") ) )  +
  ylab(NULL) +
  theme(strip.background = element_blank(),
        strip.placement = "outside") +
  xlab("Guild") + scale_fill_discrete(guide = "none")
ggsave(g, filename = "figures/Figure_3.tiff", device = "tiff", dpi = 600, bg = "white", width = 10, height = 4, compression = "lzw")
```
